---
title: "Assembling Our First Trad Rack"
author: "Joe McGirr"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    collapsed: no
    df_print: paged
    number_sections: no
    theme: paper
    toc: yes
    toc_depth: 5
    toc_float: yes
  html_notebook:
    toc: yes
    toc_depth: 5
  pdf_document:
    toc: yes
    toc_depth: '5'
editor_options:
  chunk_output_type: console
---


```{css, echo=FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
```

# Our First Trad Rack

Kimiko and I have been climbing for 5 years both indoors and outdoors. We've accumulated lots of gear for top rope, sport, and bouldering. The next thing to add to our climbing gear is a beginner's trad rack.

After taking a trad placement workshop, we decided we want the following:

	- 2 sets of cams. One set with low to mid sizes and one set with a full range that includes a couple big boys.
	- A set of nuts with a full range of sizes.
	- A set of hexes with mid to large sizes. 


# Data

We compiled a data set that includes gear from Black Diamond, Totoem, DMM, Metolius, and Wild Country based on recommendations from climbing friends.

Prices reflect manufacturer MSRP Q1 2022.

Brands and models shown below for each category.

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 3}

library(IRanges) 
library(tidyverse)

# Wong, B. Points of view: Color blindness. Nat Methods (2011).
bla <- '#000000'
blu <- '#0072b2'
grb <- '#56b4e9'
lir <- '#cc79a7'
gre <- '#009e73'
red <- '#d55e00'
org <- '#e69f00'
yel <- '#f0e442'
gry <- '#BBBBBB'

my_cols <- c(blu,red,gry,gre,org,grb,lir,gry)

gear <- read.delim("C:/Users/joseph.mcgirr/Personal/proj/climbing/placements.txt")
gear$min <- gear$min * 10
gear$max <- gear$max * 10
gear$width <- gear$max - gear$min
gear$brand_model <- paste(gear$brand, gear$model, sep = " / ")
gear$brand_model_type <- paste(gear$brand, gear$model, gear$type, sep = " / ")
gear$size <- as.character(gear$size)

gear$color_code <-  str_replace_all(gear$color_code,"bla","#BBBBBB")
gear$color_code <-  str_replace_all(gear$color_code,"blu","#0072b2")
gear$color_code <-  str_replace_all(gear$color_code,"grb","#56b4e9")
gear$color_code <-  str_replace_all(gear$color_code,"lir","#cc79a7")
gear$color_code <-  str_replace_all(gear$color_code,"gre","#009e73")
gear$color_code <-  str_replace_all(gear$color_code,"red","#d55e00")
gear$color_code <-  str_replace_all(gear$color_code,"org","#e69f00")
gear$color_code <-  str_replace_all(gear$color_code,"yel","#f0e442")
gear$color_code <-  str_replace_all(gear$color_code,"gry","#BBBBBB")
gear$color_code <-  str_replace_all(gear$color_code,"pur","#cc79a7")


cams   <- filter(gear, type == "cam") %>% pull(brand_model) %>% unique()
nuts   <- filter(gear, type == "nut") %>% pull(brand_model) %>% unique()
hexes  <- filter(gear, type == "hex") %>% pull(brand_model) %>% unique()

print("cams")
print(cams)

print("nuts")
print(nuts)

print("hexes")
print(hexes)


```

# Analyses

## Crack size coverage

### Cams

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 3}

for(cam in cams){

i_gear <- filter(gear, brand_model %in% c(cam))	

ir <- IRanges(i_gear$min,width = i_gear$width)
bins <- disjointBins(IRanges(start(ir), end(ir) + 1))
dat <- cbind(as.data.frame(ir), bin = bins)

i_gear$min <- dat$start *0.1
i_gear$max <- dat$end *0.1
i_gear$bin <- dat$bin
i_gear$width <- i_gear$width *0.1
i_gear <- i_gear %>% 
					mutate(plot_col = ifelse(type == "cam", "blu",ifelse(type == "nut","red","yel"))) %>%
					mutate(bin2 = ifelse(type == "cam", 1,ifelse(type == "nut",2,3)))


p1 <- ggplot(i_gear) + 
      geom_rect(aes(xmin = min, xmax = max,
                    ymin = bin, ymax = bin + 0.9),fill = i_gear$color_code) +
			geom_text(data = i_gear, aes(x=min+((max-min)/2), y = bin+0.5, label = size), size = 3) +
	xlab("\nrange (mm)") +
	ggtitle(paste("cam = ",cam)) +
	theme_minimal() +
	xlim(0,200) +
	theme(legend.position="none")+ 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(size=14))

print(p1)

}

```

### Nuts

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 3}

for(nut in nuts){

i_gear <- filter(gear, brand_model %in% c(nut))	

ir <- IRanges(i_gear$min,width = i_gear$width)
bins <- disjointBins(IRanges(start(ir), end(ir) + 1))
dat <- cbind(as.data.frame(ir), bin = bins)

i_gear$min <- dat$start *0.1
i_gear$max <- dat$end *0.1
i_gear$bin <- dat$bin
i_gear$width <- i_gear$width *0.1
i_gear <- i_gear %>% 
					mutate(plot_col = ifelse(type == "cam", "blu",ifelse(type == "nut","red","yel"))) %>%
					mutate(bin2 = ifelse(type == "cam", 1,ifelse(type == "nut",2,3)))


p1 <- ggplot(i_gear) + 
      geom_rect(aes(xmin = min, xmax = max,
                    ymin = bin, ymax = bin + 0.9, fill = type),fill = i_gear$color_code) +
	#scale_fill_manual(values=c(blu,red,yel)) +
	scale_fill_manual(values=c("nut" = red)) +
	xlab("\nrange (mm)") +
	ggtitle(paste("nut = ",nut)) +
	theme(legend.position="none")+ 
	theme_minimal() +
	xlim(0,40) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

print(p1)

}

```

### Hexes

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 3}

for(hex in hexes){

i_gear <- filter(gear, brand_model %in% c(hex))	

ir <- IRanges(i_gear$min,width = i_gear$width)
bins <- disjointBins(IRanges(start(ir), end(ir) + 1))
dat <- cbind(as.data.frame(ir), bin = bins)

i_gear$min <- dat$start *0.1
i_gear$max <- dat$end *0.1
i_gear$bin <- dat$bin
i_gear$width <- i_gear$width *0.1
i_gear <- i_gear %>% 
					mutate(plot_col = ifelse(type == "cam", "blu",ifelse(type == "nut","red","yel"))) %>%
					mutate(bin2 = ifelse(type == "cam", 1,ifelse(type == "nut",2,3)))


p1 <- ggplot(i_gear) + 
      geom_rect(aes(xmin = min, xmax = max,
                    ymin = bin, ymax = bin + 0.9, fill = type),fill = i_gear$color_code) +
	#scale_fill_manual(values=c(blu,red,yel)) +
	scale_fill_manual(values=c("hex" = yel)) +
	xlab("\nrange (mm)") +
	ggtitle(paste("hex = ",hex)) +
	theme_minimal() +
	xlim(0,80) +
	theme(legend.position="none")+ 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

print(p1)

}

```

## Crack size coverage with different combinations of gear

Not all combinations are shown -- just the ones that I thought would be interesting given my initial biases from researching gear.

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 3}

i_cams   <- cams[c(1,2,3)]
i_nuts   <- nuts[c(1,5)]
i_hexes  <- hexes[c(1)]

for(cam in i_cams){
	for(nut in i_nuts){
		for(hex in i_hexes){
			
i_gear <- filter(gear, brand_model %in% c(cam,nut,hex))	

ir <- IRanges(i_gear$min,width = i_gear$width)
bins <- disjointBins(IRanges(start(ir), end(ir) + 1))
dat <- cbind(as.data.frame(ir), bin = bins)

i_gear$min <- dat$start *0.1
i_gear$max <- dat$end *0.1
i_gear$bin <- dat$bin
i_gear$width <- i_gear$width *0.1
i_gear <- i_gear %>% 
					mutate(plot_col = ifelse(type == "cam", "blu",ifelse(type == "nut","red","yel"))) %>%
					mutate(bin2 = ifelse(type == "cam", 1,ifelse(type == "nut",2,3)))


p1 <- ggplot(i_gear) + 
      geom_rect(aes(xmin = min, xmax = max,
                    ymin = bin, ymax = bin + 0.9, fill = type)) +
	#scale_fill_manual(values=c(blu,red,yel)) +
	scale_fill_manual(values=c("cam" = blu, "nut" = red, "hex" = yel)) +
	xlab("\nrange (mm)") +
	ggtitle(paste("cam = ",cam,"\n","nut = ",nut,"\n","hex = ",hex,sep = "")) +
	theme_minimal() +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

print(p1)
				
		}
	}
}

#brand <- str_split(gear$brand_model," : ")[[1]][1]
#model <- str_split(gear$brand_model," : ")[[1]][2]

```

# Price

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 6}

prices <- gear %>% group_by(brand_model) %>% 
					summarize(total_price = sum(price), price_per_piece = round(sum(price)/n(),2), type = unique(type))

p1 <- filter(prices, type == "cam") %>%
			ggplot(aes(brand_model, total_price)) +
			geom_col() +
			geom_text(aes(label=total_price), position=position_dodge(width=0.9), vjust=-0.25) +
			theme_minimal() +
	    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
	    			axis.title.y=element_text(size=14))+
			ylab("total price $\n") + xlab("brand / model") + ggtitle("cams")

print(p1)

p1 <- filter(prices, type == "cam") %>%
			ggplot(aes(brand_model, price_per_piece)) +
			geom_col() +
			geom_text(aes(label=price_per_piece), position=position_dodge(width=0.9), vjust=-0.25) +
			theme_minimal() +
			theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
	    			axis.title.y=element_text(size=14))+
			ylab("price per piece $\n") + xlab("brand / model") + ggtitle("cams")

print(p1)



p1 <- filter(prices, type == "nut") %>%
			ggplot(aes(brand_model, total_price)) +
			geom_col() +
			geom_text(aes(label=total_price), position=position_dodge(width=0.9), vjust=-0.25) +
			theme_minimal() +
	    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
	    			axis.title.y=element_text(size=14))+
			ylab("total price $\n") + xlab("brand / model") + ggtitle("nuts")

print(p1)

p1 <- filter(prices, type == "nut") %>%
			ggplot(aes(brand_model, price_per_piece)) +
			geom_col() +
			geom_text(aes(label=price_per_piece), position=position_dodge(width=0.9), vjust=-0.25) +
			theme_minimal() +
	    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
	    			axis.title.y=element_text(size=14))+
			ylab("price per piece $\n") + xlab("brand / model") + ggtitle("nut")

print(p1)



p1 <- filter(prices, type == "hex") %>%
			ggplot(aes(brand_model, total_price)) +
			geom_col() +
			geom_text(aes(label=total_price), position=position_dodge(width=0.9), vjust=-0.25) +
			theme_minimal() +
	    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
	    			axis.title.y=element_text(size=14))+
			ylab("total price $\n") + xlab("brand / model") + ggtitle("hexes")

print(p1)

p1 <- filter(prices, type == "hex") %>%
			ggplot(aes(brand_model, price_per_piece)) +
			geom_col() +
			geom_text(aes(label=price_per_piece), position=position_dodge(width=0.9), vjust=-0.25) +
			theme_minimal() +
	    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
	    			axis.title.y=element_text(size=14))+
			ylab("price per piece $\n") + xlab("brand / model") + ggtitle("hexes")

print(p1)

```

# Price to Weight Ratio

```{r, message=FALSE, warning=FALSE, fig.width= 9, fig.height= 5}

n_brand_models <- gear %>% filter(type == "cam")%>% pull(brand_model) %>% unique() %>% length()

p1 <- filter(gear, type == "cam") %>%
			ggplot(aes(weight, price, color = brand_model)) +
			geom_point(size = 4) +
			geom_smooth(method = "lm", fill = "NA") +
			theme_minimal() +
			scale_color_manual(values = my_cols[c(1:n_brand_models)]) +
	    theme(axis.title.x=element_text(size=14),
	    			axis.title.y=element_text(size=14))+
			ggtitle("cams") + xlab("\nweight (g)") + ylab("price ($)\n")

print(p1)

n_brand_models <- gear %>% filter(type == "nut")%>% pull(brand_model) %>% unique() %>% length()

p1 <- filter(gear, type == "nut") %>%
			ggplot(aes(weight, price, color = brand_model)) +
			geom_point(size = 4) +
			geom_smooth(method = "lm", fill = "NA") +
			theme_minimal() +
			scale_color_manual(values = my_cols[c(1:n_brand_models)]) +
	    theme(axis.title.x=element_text(size=14),
	    			axis.title.y=element_text(size=14))+
			ggtitle("nuts") + xlab("\nweight (g)") + ylab("price ($)\n")

print(p1)

n_brand_models <- gear %>% filter(type == "nut")%>% pull(brand_model) %>% unique() %>% length()

p1 <- filter(gear, type == "hex") %>%
			ggplot(aes(weight, price, color = brand_model)) +
			geom_point(size = 4) +
			geom_smooth(method = "lm", fill = "NA") +
			theme_minimal() +
			scale_color_manual(values = my_cols[c(1:n_brand_models)]) +
	    theme(axis.title.x=element_text(size=14),
	    			axis.title.y=element_text(size=14))+
			ggtitle("hexes") + xlab("\nweight (g)") + ylab("price ($)\n")

print(p1)

```

# Strength to Weight Ratio

```{r, message=FALSE, warning=FALSE, fig.width= 9, fig.height= 5}

n_brand_models <- gear %>% filter(type == "cam")%>% pull(brand_model) %>% unique() %>% length()

p1 <- filter(gear, type == "cam") %>%
			ggplot(aes(weight, strength, color = brand_model)) +
			geom_point(size = 4) +
			geom_smooth(method = "lm", fill = "NA") +
			theme_minimal() +
			scale_color_manual(values = my_cols[c(1:n_brand_models)]) +
	    theme(axis.title.x=element_text(size=14),
	    			axis.title.y=element_text(size=14))+
			ggtitle("cams") + xlab("\nweight (g)") + ylab("strength (kN)\n")

print(p1)

n_brand_models <- gear %>% filter(type == "nut")%>% pull(brand_model) %>% unique() %>% length()

p1 <- filter(gear, type == "nut") %>%
			ggplot(aes(weight, strength, color = brand_model)) +
			geom_point(size = 4) +
			geom_smooth(method = "lm", fill = "NA") +
			theme_minimal() +
			scale_color_manual(values = my_cols[c(1:n_brand_models)]) +
	    theme(axis.title.x=element_text(size=14),
	    			axis.title.y=element_text(size=14))+
			ggtitle("nuts") + xlab("\nweight (g)") + ylab("strength (kN)\n")

print(p1)

n_brand_models <- gear %>% filter(type == "hex")%>% pull(brand_model) %>% unique() %>% length()

p1 <- filter(gear, type == "hex") %>%
			ggplot(aes(weight, strength, color = brand_model)) +
			geom_point(size = 4) +
			geom_smooth(method = "lm", fill = "NA") +
			theme_minimal() +
			scale_color_manual(values = my_cols[c(1:n_brand_models)]) +
	    theme(axis.title.x=element_text(size=14),
	    			axis.title.y=element_text(size=14))+
			ggtitle("hexes") + xlab("\nweight (g)") + ylab("strength (kN)\n")

print(p1)



```

# The winning combination

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 5}

i_cams   <- filter(gear, brand_model %in% c(cams[c(1,3)]))
i_nuts1   <- filter(gear, brand_model %in% c(nuts[c(1)]))
i_nuts   <- filter(gear, brand_model %in% c(nuts[c(5)]), size %in% c(1:5)) %>% bind_rows(i_nuts1)
i_hexes  <- filter(gear, brand_model %in% c(hexes[c(1)]))

i_gear <- bind_rows(i_cams,i_nuts,i_hexes)

ir <- IRanges(i_gear$min,width = i_gear$width)
bins <- disjointBins(IRanges(start(ir), end(ir) + 1))
dat <- cbind(as.data.frame(ir), bin = bins)

i_gear$min <- dat$start *0.1
i_gear$max <- dat$end *0.1
i_gear$bin <- dat$bin
i_gear$width <- i_gear$width *0.1
i_gear <- i_gear %>% 
					mutate(plot_col = ifelse(type == "cam", "blu",ifelse(type == "nut","red","yel"))) %>%
					mutate(bin2 = ifelse(type == "cam", 1,ifelse(type == "nut",2,3)))


p1 <- ggplot(i_gear) + 
      geom_rect(aes(xmin = min, xmax = max,
                    ymin = bin, ymax = bin + 0.9, fill = type)) +
	#scale_fill_manual(values=c(blu,red,yel)) +
	scale_fill_manual(values=c("cam" = blu, "nut" = red, "hex" = yel)) +
	xlab("\nrange (mm)") +
	ggtitle(paste("cam = WC Friends and Totems","\n",
								"nut = Wild Country Rock size 1-5 and DMM Allot Offset 7-11 ","\n",
								"hex = Wild Country Rockcentric - Sling size 5-9",sep = "")) +
	theme_minimal() +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

print(p1)
				
```

# The Damage

```{r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 5}

print(paste("cost: $",sum(i_gear$price), sep = ""))

```


